Air cargo load and route planning in pickup and delivery operations

### 2023-06-05 CAIE/paper.tex
NMO numTrials adjusted
MIPgap

			\State $f_{\pi_{TSP}} \gets SolveTour(\pi_{TSP}, L, M, C, N, method, time/(K+1))$ \label{main:f1}
			\State $answer1[scenario,surplus,method] \gets f_{\pi_{TSP}}$ \label{main:a1} \Comment{Results corresponding to $\pi_{TSP}$}
			\For {each $\pi \in S_K$} \label{main:loop1} \Comment{$\pi$ is a permutation of nodes (a tour)}
				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, time/K!)$ \Comment{$f_{\pi}$\/ is a solution to the tour $\pi$} \label{main:f2}
			\EndFor \label{main:loop2}
			\State $answer2[scenario,surplus,method] \gets \max f_{\pi}$ \label{main:a2} \Comment{Best result among all permutations}
		\EndFor

@@

@@

{\color{red}

@@

The upper bound (UB) is the objective function value of a linear programming relaxation of the problem, demonstrating that no integer viable solution can have an objective value greater (in maximization) than this.

We solved all scenarios and instances with Gurobi in a relaxed linear model without a time limit. Although the solutions are not guaranteed to be feasible, it is good for the reader to have an idea of the possible upper bound for the metaheuristic solutions.

Note that, in Table \ref{tab:20} all $Gurobi\ f$\/ values are italicized to indicate that these values may not be optimal or feasible as the decision variables are not integer.

@@

\begin{table}[H]
\centering
\caption{Methods and scenarios with $surplus =1.2$}  \label{tab:20}
\footnotesize
\begin{tabular}{c|c|ccccc|c}
\toprule
$method$& {\bf Scenarios}                 &{\bf 1}&{\bf 2} & {\bf 3}      & {\bf 4}      & {\bf 5}  & \specialcell{{\bf Normalized}\\{\bf Speed-up}}  \\
\midrule[.1pt]
{\it Gurobi}     & UB for $f$           & {\it 7.91} & {\it 11.92} & {\it 13.12} & {\it 14.29} & {\it 13.34} & -  \\ %51,84
\toprule
\multirow{2}{*}{\it TS }      & $f$     & 5.37       & 5.86        & 5.25        & 4.62        & 3.60       &   \\ %  3.60	 86
&{\bf run time (s)}                          & 659        & 361         & 251         & 105         & 86         &   \\ 
\midrule[.1pt]		
\multirow{2}{*}{\it GRASP }   & $f$     & 8.35       & 11.94       & 12.89       & 14.10       & 12.98      &   \\ % 8.35	 97	
&{\bf run time (s)}                          & 97         & 156         & 222         & 109         & 105        &   \\ 
\midrule[.1pt]
\multirow{2}{*}{\it ACO}     & $f$      & 5.25       & 5.75        & 5.06        & 4.84        &  4.20      &   \\ % 4.20	 1133
&{\bf run time (s)}                          & 313        & 326         & 303         & 256         &  453       &   \\ 
\midrule[.1pt]	
\multirow{2}{*}{\it NMO}     & $f$      & 8.62       &             &             &             &            &   \\ 
&{\bf run time (s)}                          & 901        &             &             &             &            &   \\ 
\midrule[.1pt]	
\multirow{2}{*}{\it Shims}   & $f$      & 8.43       & 12.77       & 14.47       &  16.28      &  15.47          &   \\
&{\bf run time (s)}                     & 4          &  8          &  33         &  93         &  570          &   \\

@@

\begin{table}[H]
	\centering
	\caption{Methods and scenarios with $surplus =1.5$}  \label{tab:50}
	\footnotesize
	\begin{tabular}{c|c|ccccc|c}
		\toprule
		{$method$}                        & {\bf Results}   & {\bf 1}     &  {\bf 2}    & {\bf 3}      & {\bf 4}      & {\bf 5}    & \specialcell{{\bf Normalized}\\{\bf Speed-up}} \\
		\toprule
		\multirow{2}{*}{\textbf{Gurobi}}  & $f$     & 4.756       & 8.84        & 11.94       & 16.21       &{\bf 16.47} & 0.852  \\	% 6828 
		&{\bf run time (s)}                         & 18        & 68         & 326        & 1,906      & 12,960    & 1.000  \\	 
		\midrule[.1pt]
		\multirow{2}{*}{\textbf{TS}}      & $f$     & 8.97       & 11.44       & 12.92       & 15.20       & 15.39      &  0.936 \\		
		&{\bf run time (s)}                         & 4         & 18         & 88         & 525        & 3,660     &  3.555  \\		
		\midrule[.1pt]		
		\multirow{2}{*}{\textbf{GRASP}}   & $f$     & 9.36       & 12.34       & 13.63       & 16.18       & 15.44      &  0.981 \\ 
		&{\bf run time (s)}                         & 4         & 17         & 86         & 515        & 3,564     &  3.651  \\
		\midrule[.1pt]	
		\multirow{2}{*}{\textbf{ACO}}     & $f$     & 9.36       & 12.34       & 13.63       & 16.18       & 15.44      &  0.981 \\
		&{\bf run time (s)}                         & 5         & 19         & 90         & 539        & 3,720     &  3.492  \\	
		\midrule[.1pt]	
		\multirow{2}{*}{\textbf{NMO}}     & $f$     & 9.36       & 12.33       & 13.60       & 16.16       & 15.43      &  0.979 \\ 
		&{\bf run time (s)}                         & 5         & 19         & 91         & 544        & 3,780     & 3.442  \\
		\midrule[.1pt]	
		\multirow{2}{*}{\textbf{Shims}}   & $f$     & {\bf 9.39} & {\bf 12.42} & {\bf 13.72} & {\bf 16.28} & 15.49      &  {\bf 0.986} \\	
		&{\bf run time (s)}                         & 3         & 11         & 48         & 265        & 1,720     &  {\bf 7.470}  \\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

\begin{table}[H]
	\centering
	\caption{Methods and scenarios with $surplus = 2.0$}  \label{tab:100}
	\footnotesize
	\begin{tabular}{c|c|ccccc|c}
		\toprule
		{$method$}                        & {\bf Results}    & {\bf 1}     &  {\bf 2}   & {\bf 3}    & {\bf 4}    & {\bf 5}      & \specialcell{{\bf Normalized}\\{\bf Speed-up}} \\
		\toprule
		\multirow{2}{*}{\textbf{Gurobi}}  & $f$              & 4.92      & 6.58      & 6.34       & 7.24       & 8.15      &  0.334 \\ % 9951
		                                  &{\bf run time (s)}& 29       & 106      & 493      & 2,833     & 19,380      &  1.000  \\ 
		\midrule[.1pt]
		\multirow{2}{*}{\textbf{TS}}      & $f$              & 13.72     & 17.53     & 19.28     & 21.78      & 22.35      &  0.951 \\		
		&{\bf run time (s)}                                  & 5        & 19       & 93       & 556       & 3,900     &  4.992  \\	
		\midrule[.1pt]		
		\multirow{2}{*}{\textbf{GRASP}}   & $f$        & 13.99     & 19.39     & 19.86     & 23.50      & 22.57      &  0.998 \\
		&{\bf run time (s)}                            & 6        & 21       & 97       & 571       & 3,960     &  4.906  \\	
		\midrule[.1pt]
		\multirow{2}{*}{\textbf{ACO}}     & $f$        & 13.99     & 19.39     & 19.86     & 23.50      & 22.57      &  0.998 \\
		&{\bf run time (s)}                            & 6        & 23       & 105      & 597       & 4,080     &  4.745  \\	
		\midrule[.1pt]	
		\multirow{2}{*}{\textbf{NMO}}     & $f$        & 13.99     & 19.39     & 19.85     & 23.50      & 22.56      & 0.998 \\	
		&{\bf run time (s)}                            & 9        & 27       & 120      & 681       & 4,620     & 4.184  \\	
		\midrule[.1pt]	
		\multirow{2}{*}{\textbf{Shims}}   & $f$        &{\bf 14.00} & {\bf 19.44} & {\bf 19.94} & {\bf 23.55} & {\bf 22.58} & {\bf 1.000} \\	
		&{\bf run time (s)}                            & 7        & 20       & 84       & 450       & 2,918     & {\bf 6.570}  \\
		\bottomrule
	\end{tabular}
	\normalsize
\end{table}

@@

In Tables \ref{tab:20}, \ref{tab:50} and \ref{tab:100}, $Gurobi$\/ was the slowest method and $Shims$\/ was the fastest. This was expected, as $Gurobi$\/ is a MIP solver, which seeks an optimal solution within the established constraints, usually reaching its time limit on each node, while {\it Shims} is a fast heuristic. This way, $Gurobi$\/ even demands prohibitive times (more than 5 hours in scenario 6 of Table \ref{tab:100}), while the maximum time required by $Shims$\/ did not reach one hour.

Considering the {\bf Normalized}\/ values, in Table \ref{tab:20} $Gurobi$\/ was the best ($0.997$), followed closely by $Shims$\/ ($0.939$). In Tables \ref{tab:50} and \ref{tab:100}, $Shims$ was the best ($0.986$\/ and $1.000$). As expected, the performance of $Gurobi$\/ drops sharply as the volume of shipments at the nodes increases: $0.852$\/ in Table \ref{tab:50} and $0.334$\/ in Table \ref{tab:100}.

>>>>

			\State {\color{blue} $f_{\pi_{dir}} \gets SolveTour(\pi_{dir}, L, M, C, N, method, time)$ } \label{main:fdirect}

		    \State {\color{blue} $f_{\pi_{rev}} \gets SolveTour(\pi_{rev}, L, M, C, N, method, time)$ } \label{main:freverse}

			\State {\color{blue} $answer1[scenario,surplus,method] \gets max( f_{\pi_{dir}}, f_{\pi_{rev}}  )$ } \label{main:a1}

			\For {each $\pi \in S_K$} \label{main:loop1} \Comment{$\pi$ is a permutation of nodes (a tour)}

				\State $f_{\pi} \gets SolveTour(\pi, L, M, C, N, method, time/(K+1)!)$ \Comment{$f_{\pi}$\/ is a solution to the tour $\pi$} \label{main:f2}

			\EndFor \label{main:loop2}
			\State $answer2[scenario,surplus,method] \gets \max f_{\pi}$ \label{main:a2} \Comment{Best result among all permutations}
		\EndFor

@@

{\color{blue} In line \ref{main:fdirect}, $SolveTour$ solves the lowest-cost tour, and in line 8, its reverse. }

@@

@@

{\color{red}
Let $X_{ij}^k$\/ and $Y_{iq}^k$\/ be CONTINUOUS variables, where $1 \leq i \leq m$, $1 \leq j \leq n_k$, $1 \leq q \leq m_k$\/ and $0 \leq k \leq K$.
}
The upper bound (UB) is the objective function value of a linear programming relaxation of the problem, demonstrating that no integer viable solution can have an objective value greater (in maximization) than this.

We solved all scenarios and instances with Gurobi in a {\it MIPGap} parameter equal to 1\% to control the minimal quality of the returned solution, without a time limit.

Note that, in Table \ref{tab:20} all $Gurobi\ f$\/ values are italicized to indicate that these values are at most 1\% distant from the optimal solution.

@@

\begin{table}[H]
\centering
\caption{Methods and scenarios with $surplus =1.2$ and the least cost tours criteria}  \label{tab:20leastcost}
\footnotesize
\begin{tabular}{c|c|cccc}
\toprule
$method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4}  \\
\toprule
\multirow{2}{*}{\it Gurobi* } & $f$ & 7.90    & 11.46   & 12.82   & 13.40     \\
&               {\bf run time (s)}  & 10      & 8       & 11      & 11      \\ 
\midrule[.1pt]	
\multirow{2}{*}{\it Shims}    & $f$ & 7.83    & 11.34   & 12.77   & 13.20    \\
&         {\bf run time (s)}        & 4       & 5       &  6      & 7        \\ 
\bottomrule
\multicolumn{6}{c}{*integer relaxed variables}
\end{tabular}
\normalsize
\end{table}

\begin{table}[H]
\centering
\caption{Methods and scenarios with $surplus =1.2$}  \label{tab:20}
\footnotesize
\begin{tabular}{c|c|ccccc|c}
\toprule
$method$& {\bf Scenarios}                 &{\bf 1}&{\bf 2} & {\bf 3}      & {\bf 4}      & {\bf 5}  & \specialcell{{\bf Normalized}\\{\bf Speed-up}}  \\
\midrule[.1pt]
{\it Gurobi}     & UB for $f$           & {\it 8.91} & {\it 12.92} & {\it 15.12} & {\it 16.29} & {\it 15.34} & -  \\ %51,84
\toprule

\midrule[.1pt]	
\multirow{2}{*}{\it Shims}   & $f$      & 8.43       & 12.77       & 14.47       & 16.28       & 15.15          &   \\
&{\bf run time (s)}                     & 4          &  8          &  33         & 93          & 560          &   \\

@@

@@
